---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "Fanyi Zeng"
date: "02/24/22"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

Filter the Denny’s dataframe for Alaska (AK) and save the result as dn_ak. How many Denny’s locations are there in Alaska? Three.

```{r dn-Alaska}
dn <- dennys
lq <- laquinta
dn_ak <- dn %>%
  filter(state == "AK")
nrow(dn_ak)
```

Filter the La Quinta dataframe for Alaska (AK) and save the result as lq_ak. How many La Quinta locations are there in Alaska? Two.

```{r lq-Alaska}
lq_ak <- lq %>%
  filter(state == "AK")
nrow(lq_ak)
```

### Exercise 2

Next we’ll calculate the distance between all Denny’s and all La Quinta locations in Alaska. 

Let’s join the data on Denny’s and La Quinta locations in Alaska, and take a look at what it looks like:

```{r join}
dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak
```

One way of calculating the distance between any two points on the earth is to use the Haversine distance formula. This formula takes into account the fact that the earth is not flat, but instead spherical.

```{r distance}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
dn_lq_ak <- dn_lq_ak %>%
  mutate(distance = haversine(longitude.x, latitude.x, longitude.y, latitude.y))
```

Calculate the minimum distance between a Denny’s and La Quinta for each Denny’s location. To do so we group by Denny’s locations and calculate a new variable that stores the information for the minimum distance.

```{r mindist}
dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```

### Exercise 3

It seems that Denny's is close to La Quinta in Alaska. However, everything might be close together in Anchorage and Fairbanks because they are the two big cities in Alaska.

```{r AK-map}
dn_ak <- dn_ak %>%
  mutate(establishment = "Denny's")
lq_ak <- lq_ak %>%
  mutate(establishment = "La Quinta")
dn_lq_akm <- bind_rows(dn_ak, lq_ak)
dn_lq_akm %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.1)) +
  geom_point() +
  labs(title = "Locations of Denny's and La Quinta in Alaska", subtitle = "By latitude and longitude")
```

### Exercise 4

Let's plot the locations for both establishments in the NC and TX. We can see that not every La Quinta has a Denny's nearby in NC. However, it seems to hold for TX.

```{r NC-TX}
dn <- dn %>%
  mutate(establishment = "Denny's")
lq <- lq %>%
  mutate(establishment = "La Quinta")
dn_lq <- bind_rows(dn, lq)
dn_lq %>%
  filter(state == "NC") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.5)) +
  geom_point() +
  labs(title = "Locations of Denny's and La Quinta in NC", subtitle = " By latitude and longitude")
dn_lq %>%
  filter(state == "TX") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.1)) +
  geom_point() +
  labs(title = "Locations of Denny's and La Quinta in TX", subtitle = "By latitude and longitude")
```

### Exercise 5

It seems to hold for most southern states! The degree of overlap is high in the maps below.

```{r South}
dn_lq %>%
  filter(state == "CA") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.1)) +
  geom_point() +
  labs(title = "Locations of Denny's and La Quinta in CA", subtitle = "By latitude and longitude")
dn_lq %>%
  filter(state == "FL") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.1)) +
  geom_point() +
  labs(title = "Locations of Denny's and La Quinta in FL", subtitle = "By latitude and longitude")
dn_lq %>%
  filter(state == "GA") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.1)) +
  geom_point() +
  labs(title = "Locations of Denny's and La Quinta in GA", subtitle = "By latitude and longitude")
dn_lq %>%
  filter(state == "AZ") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.1)) +
  geom_point() +
  labs(title = "Locations of Denny's and La Quinta in AZ", subtitle = "By latitude and longitude")
dn_lq %>%
  filter(state == "NM") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.1)) +
  geom_point() +
  labs(title = "Locations of Denny's and La Quinta in NM", subtitle = "By latitude and longitude")
dn_lq %>%
  filter(state == "AL") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.5)) +
  geom_point() +
  labs(title = "Locations of Denny's and La Quinta in AL", subtitle = " By latitude and longitude")
dn_lq %>%
  filter(state == "LA") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.5)) +
  geom_point() +
  labs(title = "Locations of Denny's and La Quinta in LA", subtitle = " By latitude and longitude")
dn_lq %>%
  filter(state == "MS") %>%
  ggplot(data, mapping = aes(x = longitude, y = latitude, color = establishment, alpha = 0.5)) +
  geom_point() +
  labs(title = "Locations of Denny's and La Quinta in MS", subtitle = " By latitude and longitude")
```